---
title: "presentation SC april 2019 :  benchmarks outputs"
author: "Thomas Goossens"
date: "4/23/2019"
always_allow_html: yes
output:
  revealjs::revealjs_presentation:
    css:
      - "~/Documents/code/pokyah/craw-reveal-template/theme/craw.css"
    center: no
    highlight: pygments
    incremental: yes
    self_contained: true
    slide_level: 2
    theme: default
    transition: slide
    fig_caption: false
    reveal_options:
      previewLinks: false
  md_document:
    toc: no
    toc_depth: 6
    variant: markdown_github
  word_document:
    toc: no
    toc_depth: '6'
  pdf_document: default
  odt_document:
    fig_height: 5
    fig_width: 7
  html_document:
    theme: default
    toc: yes
    toc_depth: 6
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
devtools::load_all()
library(dplyr)
library(sf)
```

---
title: "presentation SC april 2019 : The agrometeoR package"
author: "Thomas Goossens"
date: "4/23/2019"
always_allow_html: yes
output:
  revealjs::revealjs_presentation:
    css:
      - "~/Documents/code/pokyah/craw-reveal-template/theme/craw.css"
    center: no
    highlight: pygments
    incremental: yes
    self_contained: true
    slide_level: 2
    theme: default
    transition: slide
    fig_caption: false
    reveal_options:
      previewLinks: false
  md_document:
    toc: no
    toc_depth: 6
    variant: markdown_github
  word_document:
    toc: no
    toc_depth: '6'
  pdf_document: default
  odt_document:
    fig_height: 5
    fig_width: 7
  html_document:
    theme: default
    toc: yes
    toc_depth: 6
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---



# Defining the best spatialization method

## idea : test various "explorative constructions"

1 EC = 1 unique combination of 

* multiple learners
* multiple potential hyper-parameters
* multiple explanatory variables
* multiple data sources (stations) : test the interest of networks integration

## Table of EC

```{r explorative-constructions, echo = FALSE}
explorative_constructions = read.csv2("./explorative_constructions.csv", sep = ";") %>%
  dplyr::mutate_all(.funs = as.factor)
knitr::kable(head(explorative_constructions))
```

## Why EC

* deliver data with metadata
* which best EC among which tested EC's
* easy shortcut name to precisely know what we are talking about

## Testing EC's

* Conduction of batches of benchmarks experiments

# Benchmarking of EC's

## Function to conduct benchmark

The agrometeoR package offers a function for this : `makeBatchOfBenchExp`
```{r batchbenchmark, eval = FALSE }
# example
bmrsResults = makeBatchOfBenchExp(
  tasks = tasks,
  learners = agrometeorLearners,
  measures = list(rmse, mae, mse),
  keep.pred = TRUE,
  models = FALSE,
  grouping = 100,
  level = "mlr.benchmark",
  resamplings = "LOO",
  cpus = 4
) 
```

## benchmark conduction difficulties 

* memory intensive => split the results in chunks written to disks to free RAM
* paralellel computing to speed up the computations.
* run on dedicated server using `bash` in a tmux `session`
* use scp to retrieve the results from server
* hourly 2 years 5 learners => 24h of computation

## Conducted benchmarks : 

01/01/2016 -> 31/12/2017, 5 learners (KED, OK, NN1, IDW, MultiReg), with elevation as explanatory for KED, OK, and multiple linear reg.

* daily tsa max pameseb
* daily tsa max pameseb + IRM
* hourly tsa pameseb
* hourly tsa pameseb + IRM

## summary statistics of hourly tsa Pameseb : RMSE

```{r, echo = FALSE}
hourly_tsa_pameseb_rmse = read.csv2("~/Desktop/sc_april_2019/tsa hourly Pameseb summary stats rmse.csv", sep = ",")
knitr::kable(hourly_tsa_pameseb_rmse)
```

## summary statistics of hourly tsa Pameseb & IRM : RMSE

```{r, echo = FALSE}
hourly_tsa_pameseb_IRM_rmse = read.csv2("~/Desktop/sc_april_2019/tsa hourly Pameseb_IRM summary stats rmse.csv", sep = ",")
knitr::kable(hourly_tsa_pameseb_IRM_rmse)
```

## summary statistics of hourly tsa Pameseb : Residuals
```{r, echo = FALSE}
hourly_tsa_pameseb_residuals = read.csv2("~/Desktop/sc_april_2019/tsa hourly Pameseb summary stats residuals.csv", sep = ",")
knitr::kable(hourly_tsa_pameseb_residuals)
```

## summary statistics of hourly tsa Pameseb & IRM : Residuals
```{r, echo = FALSE}
hourly_tsa_pameseb__IRM_residuals = read.csv2("~/Desktop/sc_april_2019/tsa hourly Pameseb_IRM summary stats residuals.csv", sep = ",")
knitr::kable(hourly_tsa_pameseb__IRM_residuals)
```

## summary statistics of daily tsa max Pameseb : RMSE

```{r, echo = FALSE}
daily_tsamax_pameseb_rmse = read.csv2("~/Desktop/sc_april_2019/tsamax daily Pameseb summary stats rmse.csv", sep = ",")
knitr::kable(daily_tsamax_pameseb_rmse)
```

## summary statistics of daily tsa max Pameseb & IRM : RMSE

```{r, echo = FALSE}
daily_tsamax_pameseb_IRM_rmse = read.csv2("~/Desktop/sc_april_2019/tsamax daily Pameseb_IRM summary stats rmse.csv", sep = ",")
knitr::kable(daily_tsamax_pameseb_IRM_rmse)
```

## summary statistics of daily tsa max Pameseb : Residuals
```{r, echo = FALSE}
daily_tsamax_pameseb_residuals = read.csv2("~/Desktop/sc_april_2019/tsamax daily Pameseb summary stats residuals.csv", sep = ",")
knitr::kable(daily_tsamax_pameseb_residuals)
```

## summary statistics of daily tsa max Pameseb & IRM : Residuals
```{r, echo = FALSE}
daily_tsamax_pameseb_IRM_residuals = read.csv2("~/Desktop/sc_april_2019/tsamax daily Pameseb_IRM summary stats residuals.csv", sep = ",")
knitr::kable(daily_tsamax_pameseb_IRM_residuals)
```
